\section{Introduction}

Distributed systems face a fundamental coordination challenge: clocks on independent nodes inevitably diverge, yet many algorithms require agreement on time. \textit{Skew} refers to the instantaneous offset between two clocks---at time $t$, if clock $C_A$ reads $T_A$ and clock $C_B$ reads $T_B$, the skew is $|T_A - T_B|$. \textit{Drift} refers to the rate at which clocks diverge due to oscillator frequency differences, causing skew to grow linearly: $\text{skew}(t) = \text{skew}(t_0) + |\rho_A - \rho_B| \cdot (t - t_0)$. Clock synchronization protocols periodically measure and correct for skew and drift, while bounded uncertainty accounts for residual errors.

The precision gap between NTP and PTP spans three orders of magnitude. NTP~\cite{ntp} achieves 1--10\,ms accuracy on local networks through software timestamping, degrading to 10--100\,ms across the internet. NTP's fundamental limitations---software-induced jitter from OS scheduling, undetectable network asymmetry, and sparse synchronization intervals---stem from its design philosophy of ubiquitous deployment without infrastructure control. PTP~\cite{ptp} achieves sub-microsecond accuracy through hardware timestamping at the physical layer, but requires PTP-capable NICs, transparent clock switches, and GPS-disciplined grandmaster clocks~\cite{ptp_hardware_costs_survey,meta_ptp_time_appliances}---infrastructure costing \$50,000+ per datacenter that is impractical for heterogeneous deployments spanning cloud providers, edge devices, and multi-site configurations.

Between these extremes lies a vast class of applications requiring better-than-millisecond precision but lacking resources for PTP. Data streaming engines need window boundary precision of hundreds of microseconds. Distributed databases need bounded uncertainty for consistency protocols across heterogeneous cloud providers. Multi-region systems require temporal coordination across geographic sites.

Google's TrueTime~\cite{spanner_osdi2012_truetime_gps_atomic} demonstrated the transformative power of \textit{bounded uncertainty}: knowing not just ``my time is $T$'' but ``my time is $T \pm \epsilon$ with high confidence.'' With 1--7\,ms uncertainty bounds, Spanner achieves external consistency without sub-microsecond precision---what matters is that all nodes agree on bounds rather than exact values. This insight---that bounded uncertainty enables strong guarantees even with moderate absolute accuracy---motivates our approach.

The key abstraction is: events at times $t_1 \pm \epsilon_1$ and $t_2 \pm \epsilon_2$ are correctly ordered when $t_1 + \epsilon_1 < t_2 - \epsilon_2$. Systems providing accurate timestamps without uncertainty bounds force conservative worst-case assumptions, while those quantifying uncertainty enable algorithms to balance wait times against required confidence levels.

Recent advances in zero-shot time series foundation models present an opportunity to bridge the NTP-PTP gap through \textit{software-defined synchronization}, paralleling how Software-Defined Networking~\cite{sdn_openflow_mckeown_sigcomm} replaced specialized routing hardware with programmable control planes. Foundation models including Chronos~\cite{Chronos}, TimesFM~\cite{TimesFM}, and Moirai~\cite{MOIRAI} produce \textit{probabilistic forecasts} through multi-quantile outputs---predicting not just ``clock will read $X$ at time $T$'' but ``clock will read $X \pm Y$ with 90\% confidence''---providing exactly the bounded uncertainty that distributed algorithms require. Zero-shot capability enables deployment across heterogeneous hardware without per-clock training.

We present ChronoTick, a predictive clock synchronization system that provides TrueTime-like bounded uncertainty on commodity NTP infrastructure. Our contributions:

\begin{enumerate}
    \item \textbf{Sub-millisecond precision}: 0.604\,ms MAE across heterogeneous platforms, representing 2.5--3$\times$ improvement over commodity NTP, through a dual-model architecture combining short-horizon and long-horizon foundation model forecasting.
    \item \textbf{Calibrated uncertainty bounds}: 94.9\% coverage through quantile-based uncertainty extraction, meeting TrueTime-equivalent probabilistic guarantees without GPS or atomic clocks.
    \item \textbf{Zero-shot cross-platform generalization}: Immediate deployment across consumer workstations, cloud servers, and HPC clusters without per-device calibration or fine-tuning.
    \item \textbf{Production-ready operation}: Sustained 10+ hour deployments with microsecond-scale client access latency (2\,$\mu$s), defensive mechanisms for measurement anomalies, and self-healing retrospective correction.
\end{enumerate}
